\section{Motes v. Discarded Phones}
\label{sec-comparison}

\input{./figures/devicetable.tex}

As part of determining whether discarded smartphones can replace sensor motes
for some sensor networking applications, we perform an attribute-by-attribute
comparison of the two options. We compare the Epic mote, a common sensor node
platform, with one discarded phone, the Samsung Nexus~S~4G  model that was
used for our preliminary feasibility study. Table~\ref{table-comparison}
presents numbers used in the discussion below, drawn from both datasheets and
experiments. As an additional comparison point, we also include numbers for
the Raspberry Pi Model B, since this is a popular, cheap and powerful
single-board computer. Our comparison considers the multiple aspects involved
in deploying sensor networks, including cost, difficulty, deployment effort,
servicing, maintainability, and power provisioning.

\subsection{Cost and Availability}

We would expect the rapid turnover and high production volumes for consumer
devices like smartphones to cause their prices to start low and fall quickly,
and our data shows that this is the case. The Samsung Nexus~S was released in
2010 at \$529 unlocked, but only three years later Sprint offers customers
\$22 as a trade-in value for a returned device in good condition, making it
cheaper than both the Raspberry~Pi Model~B (\$35) and the Epic~Mote (\$69),
despite offering many more features as detailed below.

While both the Raspberry~Pi and Epic can be purchased new in unlimited
quantities, one concern about the use of discarded devices is availability.
However, as stated earlier the EPA estimates that 141~million mobile devices
were discarded in 2009. In several years, even if only half are Android
devices and only 10\% of those are in working condition, that still leaves
millions available for reuse. \textbf{Advantage: discarded smartphone.}

\subsection{Packaging and Human Interface}

Both the Epic~Mote and the Raspberry~Pi are shipped as bare circuit boards
and lack any human interface. In contrast, the discarded Nexus~S comes
packaged in plastic and features a familiar touch-screen interface. Even if
they are only used during deployment debugging or maintenance, a screen
provides a powerful maintainability advantage over sensor nodes, which must
communicate by flashing their LEDs. \textbf{Advantage: discarded smartphone.}

\subsection{Sensors and Sensor Interface}

The Epic~Mote is designed as a sensor platform, and so includes multiple
sensor interfaces while integrating no on-board sensors. The Raspberry~Pi is
not designed for sensing applications but still exposes GPIO ports and a USB
interface allowing external sensors to be attached.

On the Nexus~S and other discarded phones the situation is more complicated.
On the plus side, many phones include multiple integrated sensors, although
the sensor suite is designed around smartphone and mobile computing
applications and its composition varies across devices. In addition, the
increasingly-ubiquitous $\mu$USB port creates the possibility of adding
additional sensors, or the even more ubiquitous audio jack can be
``hijacked''~\cite{hijack-dev10} for this purpose.

Unfortunately, our experience is that many smartphone USB controllers are not
designed for this purpose. If they can enter USB host mode at all, their
power consumption when acting as a host is prohibitive, possibly because they
are expecting to receive power when the USB cable is connected. We are
investigating this problem in more detail and hope to be able to develop
software solutions on devices where the hardware has the necessary
capabilities.

An alternative approach to extending a device's default sensor suite is to
use a tiered approach. Instead of connecting a single additional sensor via
USB, we could connect multiple sensors to an additional piece of hardware
that is in turn connected to the Android device. In practice, this can be
accomplished using the Android Open Accessory protocol (AOA) which provides a
way to integrate custom hardware accessories with Android devices over USB. A
suitable candidate for a daughterboard in this tiered approach is the Arduino
ADK board. There are a wide range of sensors available for the Arduino
platform such as temperature, humidity, light, and water sensors. Using the
AOA protocol, it is possible to access data from these sensors on an Android
device over USB.

\begin{comment}

% 16 Jan 2014 : SDH : Removed to create space.

Using this Turducken-esque~\cite{sorber2005turducken} tiered approach, we
should also be able to increase energy efficiency. By offloading sensing
tasks to the Arduino, we can reduce the devices's duty cycle, only needing to
periodically wake up to collect the samples from the Arduino.  Furthermore,
the Arduino should be much more efficient at sensing as the task better fits
its design. Using the discarded smartphone as the top level in this tier
still yields the advantage of being able to process and transmit this data on
the more-powerful and better-connected smartphone.
\end{comment}

In summary, while smartphones were not designed to be sensor platforms, they
include built-in sensors and can be extended to control a variety of others,
albeit with the overheads of including a second device. In contrast, motes
lack the built-in sensors but are designed to integrate with external ones
easily. \textbf{Advantage: even.}

\subsection{Programming Environment}

Both the TinyOS~\cite{tinyos} and Contiki~\cite{contiki} sensor node
programming frameworks are notoriously difficult to learn and use, a
challenge that the sensor network research community has been trying to
address for years. In contrast, there are over 1~million applications hosted
on the Google Play store, evidence of a large and growing Android developer
community built on pre-existing familiarity with the Java programming
language. We anticipate that the capabilities of these developers can be
harnessed in developing applications that reuse discarded phones as sensors
and in other ways. \textbf{Advantage: discarded smartphone.}

\subsection{Capabilities}

\sloppypar{When comparing the core device capabilities, the discarded phone
has the clear advantage, with a three orders-of-magnitude faster processor,
four orders-of-magnitude more memory, and an one order-of-magnitude more
storage. The discarded phone actually has more memory than the Epic has Flash
storage, allowing it to cache information in RAM and avoid the energy
overhead of writing to or reading from Flash. While Flash has the benefit of
persisting across failures, motes tiny memory sizes typically cause data that
could be storage in RAM to be move to Flash simply due to lack of space. The
Raspberry Pi's core specifications are similar to the discarded smartphone,
with the exception that storage is not included and must be added separately
at additional cost. \textbf{Advantage: discarded smartphone.}}

\subsection{Connectivity}

With their multiple connections to widely-deployed networks, connectivity is
another important area where the discarded Nexus~S~4G has a distinct
advantage. The amount of the world not served by cellular, mobile data, and
Wifi networks is shrinking, meaning that a smartphones can be deployed almost
anywhere without provisioning network infrastructure. In contrast, the Epic
mote has only an 802.15.4 low power, low bitrate radio, relying on a standard
that has yet to be widely deployed. While some have called for a Wifi-scale
deployment of 802.15.4~\cite{leaves-ipsn12}, we consider it unlikely that
anyone will build a new networking infrastructure at that scale to support
low-power sensor nodes. In addition, when data aggregation is possible Wifi
becomes more much more energy-efficient than
802.15.4~\cite{wirelessenergyconsumption-url}.

The ad-hoc routing and networking required when nodes lack direct connections
to the infrastructure takes a severe toll on the real lifetime of deployed
sensor networks, with an energy breakdown compiled by Klues et.
al~\cite{icem-sosp07} of a sensor node showing that periodic idle listening
reduced lifetime by an order-of-magnitude. Specialized networks also
complicate deployment logistics, with several experience
papers~\cite{volcano-osdi06,hitchhikers-sensys08} documenting the
difficulties establishing and maintaining 802.15.4 networks and bridges with
the outside world. As an anecdote, one of the authors spent several days of a
short deployment working to establish a point-to-point serial link between a
volcano-monitoring sensor network and a monitoring site located 8~km
away~\cite{lance-sensys08}. At one point, as he and a colleague were
struggling high on the volcano to diagnose a networking problem, the
colleague's phone rang with a call from movers in Berkeley, CA, some
\num{6000}~km away.

\begin{comment}

% 16 Jan 2014 : SDH : Space

While attempting to save energy, smartphones can also harness the powered
nature of the wireless infrastructure they utilize. Wifi Power Saving Mode
(PSM), for example, allows high-power routers to buffer incoming packets to
allow low-power clients to save energy. Cellular networks also buffer data,
such as text messages, for clients, and here the buffering timeouts are much
longer. A sensor network built from discarded smartphones that wanted to take
a Koala-like~\cite{koala-ipsn08} approach to providing pull-based data
collection could use text messages to awake the network. Discarded phones
could be aggressively duty-cycled, relying on the willingness of the
networking infrastructure to hold incoming messages across long sleep
intervals.

\end{comment}

Utilizing these ubiquitous networks, however, requires carriers willing to
tailor cost-effective plans allowing sharing across large numbers of nodes.
In our conversations with Sprint, they expressed interest preliminary
willingness in offering such plans. Interestingly, the ``all you can eat''
data plans offered for years by US cellular carriers may have discouraged
this use model, since without a way to meter data carriers may fear that
larger numbers of devices mean more data. Now multiple carriers offer metered
data plans that can be shared across multiple devices, indicating that they
have technology in place to implement this option.

Overall, by avoiding ad-hoc networking and tapping into the most
widely-deployed networking infrastructures available, the discarded phone is
better-connected than the mote. \textbf{Advantage: discarded smartphone.}

\subsection{Power Consumption}

As the table shows, the idle power consumption of the mote is two
orders-of-magnitude lower than the discarded phone, unsurprising given that
motes are designed to be energy-efficient. The Raspberry Pi's much higher
power consumption reflects its intended use as a powered device.
\textbf{Advantage: mote.}
